Reduced GW invariants from cuspidal curves

### 2017-12-21 draft.tex
current dropbox version

\begin{lem}\label{lem:classificationsing}(Classification of CM genus $1$ singularities)
The Cohen-Macaulay genus $1$ singularities with $m$ branches are all and only obtained by gluing a genus $0$ singularity with $k$ branches with a Smyth's $m-k$-elliptic fold. We call such a singularity a \textbf{(k,m)-type} genus $1$ singularity.
Notice that $k$ may be $0$ (i.e. a point) or $1$ (i.e. a smooth rational curve).
\end{lem}
\begin{proof}
The proof follows the technique already used by Smyth \cite[Appendix~A]{SMY1} to classify the Goreinstein genus $1$ singularities. The question of classifying singularities in local analytic, so as in \cite[Appendix~A]{SMY1} we use the following notation:
$R$ denote the analytic local ring of the singularity; 
$\widetilde{R}=k[[t_1]\oplus\dots\oplus k[[t_m]]$ its integral closure; $\mathfrak{m}_R$ is the maximal ideal in $R$  and $\mathfrak{m}_{\widetilde{R}}$ the maximal ideal in $\widetilde{R}.$

@@

\noindent{For $m\geq 3$} three possibilities:
\begin{enumerate}
\item At least two of the $a_i$ are non zero.
\item Only one of the $a_i$ is non zero
\item All the $a_i$ are zero
\end{enumerate}
\noindent{For $m=2$} two possibilities
\item $a_1$ is non zero, then we have the tacnode;
\item  $a_1$ is zero
\noindent{For $m= 1$} As in Smyth's we prove that in this  case $\mathfrak{m}^2_{\widetilde{R}}=\mathfrak{m}_{R}$ which means the singularity is a cusp. 
\end{proof}
\begin{lem}
Every Cohen-Macaulay genus $1$ singularity is smoothable.
\end{lem}
\begin{proof}
We exhibit explicit smoothings for the above singularities.
Start with two families $\cC^0\to T$ and $\cC^1\to T$ over a DVR scheme, respectively of smooth rational curves and smooth elliptic curves, with $T$ a DVR scheme, e.g. consider two trivial families.

@@

Viceversa, let us suppose that $f\colon C\to \PP^r$ factors through $\hat{f}\colon \hat{C}\to\PP^r$  with $\hat{C} $ a genus $1$ singularity.

@@

Since all the curve we are considering are Cohen-Maculay, they have a dualising pure sheaf $\omega_{\hat{C}}$ such that for any line bundle $\hat{L}$
\[H^1(\hat{C},\hat{L})=(H^0(\hat{C},\hat{L}^{-1}\otimes\omega_{\hat{C}}))^{\vee}.\]
It is well know \cite[IV, \S~3]{serre} that  $\omega_{\hat{C}}$ can be described as the subsheaf of $\nu_* \omega_{\overline{C}}\otimes K(\overline{C})$ locally satisfying:
\[\omega\in \nu_* \omega_{\overline{C}}\otimes K(\overline{C})(U)\;\;\;\text{s.t.}\;\;\sum_{q\in \nu^{-1}(0)}\rm{Res}(\nu^*f\omega) =0\;\;\;\forall \,f\in\,\OO_C(U).\]
Let $\hat{C}$ be a genus $1$ curve with only one singularity which is a genus $1$ singularity of $(k,m)$-type
 Using the explicit local description of genus $1$ singularities that we worked out in Lemma~\ref{lem:classificationsing}, we see that 
\[\omega_C\hookrightarrow \nu_*\omega_{\overline{C}}(p_1+\dots p_k+2 p_{k+1}\dots+2p_m)=\nu_*\left(\OO_{\PP^1}(-1)\oplus\dots\oplus\OO_{\PP^1}\oplus\dots \OO_{\PP^1}\right)\]
and the kernel of the inclusion is supported on the singular point. We then have an inclusion
\[H^0(\hat{C},\hat{L}^{-1}\otimes\omega_{\hat{C}})\hookrightarrow H^0(\nu_*\left(\OO_{\PP^1}(-1)\oplus\dots\oplus\OO_{\PP^1}\oplus\dots \OO_{\PP^1}\right)\otimes\hat{L})^{-1})=0;\]
where the last equality follows from the fact that $\hat{L}){-1}$ has degree less or equal to $0$ on every irreducible component of $\hat{C}.$

We conclude showing how to smooth $f\colon C\to \PP^r$ given $\hat{F}\colon\hat{C}\to\PP^r.$

@@

\section{From $p$-fields to the quintic three-fold}\label{section:p-fields}

We shall introduce the moduli space $\Mone{1}{\PP^4}{d}^p$ of $1$-stable curves with $p$-fields, endow it with a $0$-dimensional virtual class, and show that its degree coincides with the genus $1$ cuspidal invariants of the quintic three-fold $X$ up to a sign. This is a word-by-word repetition of the arguments in \cite{CLpfields}, once noticed that they carry over to the situation where we work over a family of Gorenstein (not necessarily nodal) curves. Our aim in this section is to improve the legibility of our paper by providing the non-expert reader with a r\'{e}sum\'{e} of some of the key ideas contained in \cite{CLpfields}; it should otherwise be skipped.

@@

Finally it will be useful to introduce
\bcd
\XP\ar[r]\ar[d]\ar[dr,phantom,"\Box"] & \mathfrak{Pic}_1^{(1)}\ar[d] \\
\X\ar[r] & \MM_1^{(1)}
\ecd
so that objects of $\XP(S)$ are
\bcd
\cC\ar[rr,"\phi"]\ar[dr,"\pi"] & & \hat\cC\ar[dl,"\hat\pi"] & \text{and} & \hL\in\Pic_{\hat \cC/S}(S) \\
& S & & &
\ecd
Notice that the map $\XP\to\mathfrak{Pic}$ that associates to an object as above $(\cC\to S,\phi^*\hL)$ is highly non-trivial on the locus of elliptic tails/cusps, since
\begin{itemize}
\item line bundles of the same degree on the cusp are identified (there is a $\mathbb G_a$ wealth of them), and
\item only line bundles that restrict to effective line bundles on the elliptic tail are in the image of this morphism.
\end{itemize}
\end{comment}
We conclude this brief section remarking that the blow up procedure does not affect the invariants:
\begin{lem}
We have the identity:
\[\deg\virloc{\tZp}=\deg\virloc{\Zp}\]
\end{lem}
Compare with \cite[Proposition~2.5]{CL}.

\section{Splitting the cone and proof of the Main Theorem}\label{section:main_proof}
We are finally in a position to study the cone $\mathfrak{C}_{\tZp/\tXP}.$ This is essentially going to be a word-by-word repetition of the arguments in \cite{CLpfields}.
\begin{lem}\label{lem:open_cones}
Let $\tilde{\rho}\colon \tZp\to\tXP$ be the natural map and $\mathbb {E}_{\tZp/\tXP}=\R\hat\pi_*(\mathcal L^{\oplus 5}\oplus\mathcal P)$  the relative perfect obstruction theory for $\tilde{\rho},$ then $\mathfrak{C}_{\tZp/\tXP}$ has the following properties:
\begin{enumerate}
\item Its restriction to  $\tZ^{p,\circ}= \tZ^{p,\rm{main}}\setminus \bigcup_{k\geq 2}D^k \tZp$ can be described as the zero section of $h^1/h^0(\mathbb {E}_{\tZp/\tXP})\rvert_{\tZ^{p,\circ}}$
\item Its restriction to $\tZ^{p,gst,\circ}=\tZp - \tZ^{p,\rm{main}}$  is a rank $2$ subbundle stack of $h^1/h^0(\mathbb {E}_{\tZp/\tXP})\rvert_{\tZ^{p,gst,\circ}}$
\end{enumerate}
\end{lem}
\begin{proof} Compare with \cite[Lemma 4.3]{CLpfields}
\begin{enumerate}
\item Observe that $\tZ^{p,\circ}\cong \tZ^{\circ}$ because here $H^0(\hat C, L^{\otimes -5}\otimes \omega_{\hat C})=0$. Moreover $\tZ^{\circ}$ is unobstructed on its image, which is an open of $\tXP$, because $\R^1\hat{\pi}_*\mathcal L=0$.
So the normal cone is $[\tZ^{p,\circ}/\hat{\pi}_*\mathcal L^{\oplus 5}]$, which is the zero section of $h^1/h^0(\mathbb {E}_{\tZp/\tXP})\rvert_{\tZ^{p,\circ}}=[0\oplus \R^1\hat{\pi}_*\mathcal P/\hat{\pi}_*\mathcal L^{\oplus 5}\oplus 0]$.
\item We know that $\tZ^{p,gst,\circ}$ is a line bundle over $\tZ^{gst,\circ}$. From the equations \ref{rmk:eqn} we see that the latter is smooth over its image $\pazocal W$ in $\tXP$, which is the codimension $2$ locus where the minimal genus $1$ subcurve has weight $0$ and the line bundle is trivial on it. Recall that every smooth morphism $A\to B$ factors as $A\xrightarrow{\acute{e}t} B\times \Aaff^n\xrightarrow{\pr_1} B$. So we have
\bcd
\tZ^{p,gst,\circ} \ar[r,"\acute{e}t"] & \pazocal W\times\Aaff^{5d+5}\ar[r,hook]\ar[d,"q"] & \tXP\times \Aaff^{5d+5}\ar[d] \\
& \pazocal W \ar[r,hook] & \tXP
\ecd
where the bottom horizontal arrow is a codimension $2$ regular embedding. Thus 
\[\mathfrak{C}_{\tZp/\tXP}\rvert_{\tZ^{gst,\circ}}\cong \left[ q^* C_{\pazocal W/\tXP}/ \hat{\pi}_*\mathcal L^{\oplus 5}\right]\]
is a rank 2 subbundle stack of  $h^1/h^0(\mathbb {E}_{\tZp/\tXP})\rvert_{\tZ^{p,gst,\circ}}.$
\end{enumerate}
\end{proof}

>>>>

\begin{lem}(Classification of CM genus $1$ singularities)
The Cohen-Macaulay genus $1$ singularities with $m$ branches are all and only obtained by gluing a genus $0$ singularity with $k$ branches with a Smyth's $m-k$-elliptic fold.
Notice that $k$ may be $0$ (i.e. a point) or $1$ (i.e. a smooth rational curve).
\end{lem}
\begin{proof}
The proof follows the technique already used by Smyth \cite[Appendix~A]{SMY1} to classify the Goreinstein genus $1$ singularities. The question of classifying singularities in local analytic, so as in \cite[Appendix~A]{SMY1} we use the following notation:
$R$ denote the analytic local ring of the singularity; 
$\widetilde{R}=k[[t_1]\oplus\dots\oplus k[[t_m]]$ its integral closure; $\mathfrak{m}_R$ is the maximal ideal in $R$  and $\mathfrak{m}_{\widetilde{R}}$ the maximal ideal in $\widetilde{R}.$

@@

\noindent{For $m\geq 3$} three possibilities:
\begin{enumerate}
\item At least two of the $a_i$ are non zero.
\item Only one of the $a_i$ is non zero
\item All the $a_i$ are zero
\end{enumerate}
\noindent{For $m=23$} two possibilities
\item $a_1$ is non zero, then we have the tacnode;
\item  $a_1$ is zero
\noindent{For $m= 1$} As in Smyth's we prove that in this  case $\mathfrak{m}^2_{\widetilde{R}}=\mathfrak{m}_{R}$ which means the singularity is a cusp. 
\end{proof}
\begin{lem}
Every Cohen-Macaulay genus $1$ singularity is smoothable.
\end{lem}
\begin{proof}
We exhibit explicit smoothings for the above singularities.
Start with two families $\cC^0\to T$ and $\cC^1\to T$ over a DVR scheme, respectively of smooth rational curves and smooth elliptic curves, with $T$ a DVR scheme, e.g. consider two trivial families.

@@

Viceversa, let us suppose that $f\colon C\to \PP^r$ factors through $\hat{f}\colon \hat{C}\to\PP^r$  with $\hat{C} $a genus $1$ singularity.

@@

Since all the curve we are considering are Cohen-Maculay, they have a dualising pure sheaf $\omega_{\hat{C}}$ such that for any line bundle $\hat{L}$
\[H^1(\hat{C},\hat{L})=(H^0(\hat{C},\hat{L}^{-1}\otimes\omega_{\hat{C}}))^{\vee}.\]
It is well know \cite[IV, \S~3]{serre} that  $\omega_{\hat{C}}$ can be described as the subsheaf of $\nu_* \omega_{\overline{C}}\otimes K(\overline{C})$ locally satisfying:
\[\omega\in \nu_* \omega_{\overline{C}}\otimes K(\overline{C})(U)\;\;\;\text{s.t.}\;\;\sum_{q\in \nu^{-1}(0)}\rm{Res}(\nu^*f\omega) =0\;\;\;\forall \,f\in\,\OO_C(U)\]

@@

\section{From $p$-fields to the quintic three-fold}\label{section:p-fields}
We shall introduce the moduli space $\Mone{1}{\PP^4}{d}^p$ of $1$-stable curves with $p$-fields, endow it with a $0$-dimensional virtual class, and show that its degree coincides with the genus $1$ cuspidal invariants of the quintic three-fold $X$ up to a sign. This is a word-by-word repetition of the arguments in \cite{CLpfields}, once noticed that they carry over to the situation where we work over a family of Gorenstein (not necessarily nodal) curves. Our aim in this section is to improve the legibility of our paper by providing the non-expert reader with a r\'{e}sum\'{e} of some of the key ideas contained in \cite{CLpfields}; it should otherwise be skipped.

@@

Finally it will be useful to introduce
\bcd
\XP\ar[r]\ar[d]\ar[dr,phantom,"\Box"] & \mathfrak{Pic}_1^{(1)}\ar[d] \\
\X\ar[r] & \MM_1^{(1)}
\ecd
so that objects of $\XP(S)$ are
\bcd
\cC\ar[rr,"\phi"]\ar[dr,"\pi"] & & \hat\cC\ar[dl,"\hat\pi"] & \text{and} & \hL\in\Pic_{\hat \cC/S}(S) \\
& S & & &
\ecd
Notice that the map $\XP\to\mathfrak{Pic}$ that associates to an object as above $(\cC\to S,\phi^*\hL)$ is highly non-trivial on the locus of elliptic tails/cusps, since
\begin{itemize}
\item line bundles of the same degree on the cusp are identified (there is a $\mathbb G_a$ wealth of them), and
\item only line bundles that restrict to effective line bundles on the elliptic tail are in the image of this morphism.
\end{itemize}
\end{comment}
We conclude this brief section remarking that the blow up procedure does not affect the invariants:
\begin{lem}
We have the identity:
\[\deg\virloc{\tZp}=\deg\virloc{\Zp}\]
\end{lem}
Compare with \cite[Proposition~2.5]{CL}.
\section{Splitting the cone and proof of the Main Theorem}\label{section:main_proof}
We are finally in a position to study the cone $\mathfrak{C}_{\tZp/\tXP}.$ This is essentially going to be a word-by-word repetition of the arguments in \cite{CLpfields}.
\begin{lem}\label{lem:open_cones}
Let $\tilde{\rho}\colon \tZp\to\tXP$ be the natural map and $\mathbb {E}_{\tZp/\tXP}=\R\hat\pi_*(\mathcal L^{\oplus 5}\oplus\mathcal P)$  the relative perfect obstruction theory for $\tilde{\rho},$ then $\mathfrak{C}_{\tZp/\tXP}$ has the following properties:
\begin{enumerate}
\item Its restriction to  $\tZ^{p,\circ}= \tZ^{p,\rm{main}}\setminus \bigcup_{k\geq 2}D^k \tZp$ can be described as the zero section of $h^1/h^0(\mathbb {E}_{\tZp/\tXP})\rvert_{\tZ^{p,\circ}}$
\item Its restriction to $\tZ^{p,gst,\circ}=\tZp - \tZ^{p,\rm{main}}$  is a rank $2$ subbundle stack of $h^1/h^0(\mathbb {E}_{\tZp/\tXP})\rvert_{\tZ^{p,gst,\circ}}$
\end{enumerate}
\end{lem}
\begin{proof} Compare with \cite[Lemma 4.3]{CLpfields}
\begin{enumerate}
\item Observe that $\tZ^{p,\circ}\cong \tZ^{\circ}$ because here $H^0(\hat C, L^{\otimes -5}\otimes \omega_{\hat C})=0$. Moreover $\tZ^{\circ}$ is unobstructed on its image, which is an open of $\tXP$, because $\R^1\hat{\pi}_*\mathcal L=0$.
So the normal cone is $[\tZ^{p,\circ}/\hat{\pi}_*\mathcal L^{\oplus 5}]$, which is the zero section of $h^1/h^0(\mathbb {E}_{\tZp/\tXP})\rvert_{\tZ^{p,\circ}}=[0\oplus \R^1\hat{\pi}_*\mathcal P/\hat{\pi}_*\mathcal L^{\oplus 5}\oplus 0]$.
\item We know that $\tZ^{p,gst,\circ}$ is a line bundle over $\tZ^{gst,\circ}$. From the equations \ref{rmk:eqn} we see that the latter is smooth over its image $\pazocal W$ in $\tXP$, which is the codimension $2$ locus where the minimal genus $1$ subcurve has weight $0$ and the line bundle is trivial on it. Recall that every smooth morphism $A\to B$ factors as $A\xrightarrow{\acute{e}t} B\times \Aaff^n\xrightarrow{\pr_1} B$. So we have
\bcd
\tZ^{p,gst,\circ} \ar[r,"\acute{e}t"] & \pazocal W\times\Aaff^{5d+5}\ar[r,hook]\ar[d,"q"] & \tXP\times \Aaff^{5d+5}\ar[d] \\
& \pazocal W \ar[r,hook] & \tXP
\ecd
where the bottom horizontal arrow is a codimension $2$ regular embedding. Thus 
\[\mathfrak{C}_{\tZp/\tXP}\rvert_{\tZ^{gst,\circ}}\cong \left[ q^* C_{\pazocal W/\tXP}/ \hat{\pi}_*\mathcal L^{\oplus 5}\right]\]
is a rank 2 subbundle stack of  $h^1/h^0(\mathbb {E}_{\tZp/\tXP})\rvert_{\tZ^{p,gst,\circ}}.$
\end{enumerate}
\end{proof}
